Composition-climate paper

### 2015-10-28 Composition_Ecology.Rmd
Added some bib references, worked on introduction too, cleaning up the text around confounding, conflicting and contradicting effects.

Compounding effects occur when both climatic change and land use change are biased in the same direction, *e.g.*, warming climate is compounded by land use change that is biased towards warmer environments.  In the case of warming, a species may lose climatic suitability along the warmest margins of its range at the same time as land use pressure increases, hastening decline in response to warming.  This pattern is seen for butterfly species in the northern Sierra Nevada mountains of California, where warming is compounded by land use conversion at the lowest elevations [@forister2010compounded].  Confounding effects may cause little overall effect on species distributions since land use pressure falls within the central distribution for the environmental variable. Confounding effects may increase the variability of predictive models, but do little to the overall envelope.  Garry oak distributions in the Pacific Northwest may suffer from confounding effects.  The spatial extent of the species has likely changed little, but it has been excluded from the deep soil sites it formerly occupied as a result of land use conversion for urban development and agriculture [@macdougall2004defining]. Contradicting effects result in a narrower climate niche for taxa, where warming forces the distribution mean higher as populations in one tail of the climate distribution experience greater environmental pressure, and, secondarily environmentally biased land use pressure in the other end of the tail excludes the taxa from those regions.  Here we might think of downward migrations of some bird species in the Sierra Nevadas in response to shifts in precipitation [@tingley2012push], where downward migration may be limited by land use conversion in the lowlands.

In the midwestern United States we see a strong gradient of temperature and precipitation that results in two major ecotones, one from prairie/savanna to closed forest, in an approixmately west to east direction, and one from conifer dominated, sub-boreal forests to decidious forest in a north to south direction.  Goring et al. [-@goring2015preset] have shown that land use change has significantly changed the strength and structure of this ecotone, and work across this regions has consistently shown a significant imprint of human land use change on this region [@rhemtulla2009legacies;@schulte2001original;@schulte2007homogenization;@radeloff2000historical]. To date researchers have not examined how these shifts might affect the relationships between climate and the distributions of individual taxa. Given that much of the land use change within the region occurs in the south, where agricultural conversion has largely eliminated open forests [@rhemtulla2009legacies], we might expect to see that species with more southerly distributions would show greater impacts of land use change on their distributions in climate space, assuming no shift in regional climate.  Conversely, species in the north should show little change if we expect that pre-settlement trees show similar regeneration patterns (with respect to simple presence/absence) following widespread logging.

>>>>

Compounding effects occur when both climatic change and land use change are biased in the same direction, *e.g.*, warming climate is compounded by land use change that is biased towards warmer environments.  In the case of warming, a species may lose climatic suitability along the warmest margins of its range at the same time as land use pressure increases, hastening decline in response to warming.  This pattern is seen for butterfly species in the northern Sierra Nevada mountains of California, where warming is compounded by land use conversion at the lowest elevations [@forister2010compounded].  Confounding effects may cause little overall effect on species distributions since land use pressure falls within the central distribution for the environmental variable. Confounding effects may increase the variability of predictive models, but do little to the overall envelope.  Garry oak distributions in the Pacific Northwest may suffer from confounding effects.  The spatial extent of the species has likely changed little, but it has been excluded from the deep soil sites it formerly occupied as a result of land use conversion for urban development and agriculture [@macdougall2004defining;@pellatt2014environmental]. Contradicting effects result in a narrower climate niche for taxa, where warming forces the distribution mean higher as populations in one tail of the climate distribution experience greater environmental pressure, and, secondarily environmentally biased land use pressure in the other end of the tail excludes the taxa from those regions.  Here we might think of downward migrations of some bird species in the Sierra Nevadas in response to shifts in precipitation [@tingley2012push], where downward migration may be limited by land use conversion in the lowlands.

In the midwestern United States we see a strong gradient of temperature and precipitation that results in two major ecotones, one from prairie/savanna to closed forest, in an approixmately west to east direction, and one from conifer dominated, sub-boreal forests to decidious forest in a north to south direction.  Goring et al. [-@goring2015preset] have shown that land use change has significantly changed the strength and structure of this ecotone, and work across this regions has consistently shown a significant imprint of human land use change on this region [@rhemtulla2009legacies;@schulte2001original;@schulte2007homogenization;@radeloff2000historical]. To date researchers have not examined how these shifts might affect the relationships between climate and the distributions of individual taxa. Given that much of the land use change within the region occurs in the south, where agricultural conversion has largely eliminated open forests [@rhemtulla2009legacies] we might expect to see that species with more southerly distributions would show greater impacts of land use change on their distributions in climate space, assuming no shift in regional climate.  Conversely, species in the north should show little change if we expect that pre-settlement trees show similar regeneration patterns (with respect to simple presence/absence) following widespread logging.
